---
title: "DNM_power_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE )
library(statmod)
library(purrr)
library(tidyr)
library(ggplot2)
```


```{r calculate fetal brain promoter power}
# from statmod package
#power.fisher.test(p1, p2, n1, n2, alpha=0.05, nsim=100, alternative="two.sided")
# p1,p2, proportions to be compared
# n1,n2, sample sizes
# alpha, significance level
# nsim, number of data sets to simulate

# Assuming an average of 70 DNMs per family, roughly 133000 for 1900 families. 700000 for 10k families.
# Increments of 100000/70 mutations ~1500 families
args_n <- list(n1 = seq(1, 5000000, 100000), n2 = seq(1, 5000000, 100000))

promoter_df = data.frame(sample_size = seq(1, 5000000, 100000),
                 fetalbrain_promoter_power = unlist(pmap(args_n, power.fisher.test, p1 = 1806/134969, p2 = 1732/131896, nsim = 75, alternative = "greater")))
```

```{r reshape brain promoter df}
brain_promoter_long_df <-
  promoter_df %>%
  pivot_longer(!sample_size, names_to = "annotation", values_to = "power")
```

```{r plot brain promoter power}
ggplot(brain_promoter_long_df, aes(x=sample_size/70, y=power)) + 
  #geom_point(aes(color=annotation)) + 
  geom_smooth(method = "loess", se = FALSE, aes(color=annotation), span = .9) +
  geom_hline(yintercept = .8, linetype = "dashed") +
  geom_vline(xintercept = 1917) + #number of observed proband mutations
  geom_curve(aes(x = 20000, y = .15, xend = 2200, yend = .03),arrow = arrow(length = unit(0.03, "npc")), curvature = .1) +
  annotate('text', x = 28000, y = .15, label = 'This Study' ) +
  scale_x_continuous(name = "Number of Families") +
  scale_y_continuous(name = "Power", breaks = c(0,.2,.4,.6,.8,1), limits = c(0,1)) +
  theme_minimal(base_size = 12) +
  theme(axis.title.x = element_text(margin = margin(t=12, r=0, b=0, l=0)), #space between axis labels and title
        axis.title.y = element_text(margin = margin(t=0, r=10, b=0, l=0)))

```

####Fixed sample size (1917 families), and fixed proportion of variants of interest in siblings (~.023)
### For mutations in fetal brain enhancers proband (3107/134969) and sibling (3041/131896)
#### The observed proportion is a little larger in the siblings than in the probands, so power should be zero until the proband proportion is greater
```{r calculate enhancer power with fixed sibling proportion}
args_p1 <- list(p1 = seq(.015, .110, .005))

enhancer_df = data.frame(proband_proportion = seq(.015, .110, .005),
                 power = unlist(pmap(args_p1, power.fisher.test, n1 = 134969, n2 = 131896, p2 = 3041/131896, nsim = 75, alternative = "greater")))

```

```{r plot enhancer power}
ggplot(enhancer_df, aes(x=proband_proportion, y=power)) + 
  #geom_point() + 
  geom_smooth(method = "loess", se = FALSE) +
  geom_hline(yintercept =.80, linetype = "dashed") +
  geom_point(aes(x=.02302, y=.49), size = 3) + 
  annotate('text', x = .04, y = .27, label = 'This Study' ) +
  geom_segment(aes(x = .035, y =.3 , xend = .025, yend = .46),
                  arrow = arrow(length = unit(0.5, "cm")))
  
```

So, for a power of at least .80, we can say we need a proportion of around .03 variants of interest in probands.  
Thatd be around 4,050 DNMs.  
Inputting that number into the proportions, power would be:
```{r}
#power.fisher.test(n1 = 1917, n2 = 1917, p1 = 4050/134969, p2 = 3041/131896, nsim = 75)
```
  


###TURF power calculations

```{r calculate brain turf power}
args_n <- list(n1 = seq(1, 5000000, 100000), n2 = seq(1, 5000000, 100000))

brain_turf_df = data.frame(sample_size = seq(1, 5000000, 100000),
                 brain_turf_top15_power = unlist(pmap(args_n, power.fisher.test, p1 = 20088/134969, p2 = 19768/131896, nsim = 75, alternative = "greater")),
                 brain_turf_top10_power = unlist(pmap(args_n, power.fisher.test, p1 = 13471/134969, p2 = 13141/131896, nsim = 75, alternative = "greater")),
                 brain_turf_top5_power = unlist(pmap(args_n, power.fisher.test, p1 = 6828/134969, p2 = 6540/131896, nsim = 75, alternative = "greater")),
                 brain_turf_top1_power = unlist(pmap(args_n, power.fisher.test, p1 = 1353/134969, p2 = 1293/131896, nsim = 75, alternative = "greater")))


```

```{r reshape brain turf df}
brain_turf_long_df <-
  brain_turf_df %>%
  pivot_longer(!sample_size, names_to = "annotation", values_to = "power")
```

```{r plot brain turf power}
ggplot(brain_turf_long_df, aes(x=sample_size, y=power)) + 
  #geom_point(aes(color=annotation)) + 
  geom_smooth(method = "loess", se = FALSE, aes(color=annotation), span = .9) +
  geom_hline(yintercept = .8, linetype = "dashed")
```
 
```{r calculate generic turf}
args_n <- list(n1 = seq(1, 5000000, 100000), n2 = seq(1, 5000000, 100000))

generic_turf_df = data.frame(sample_size = seq(1, 5000000, 100000),
                 generic_turf_top15_power = unlist(pmap(args_n, power.fisher.test, p1 = 20625/134969, p2 = 20091/131896, nsim = 75, alternative = "greater")),
                 generic_turf_top10_power = unlist(pmap(args_n, power.fisher.test, p1 = 13672/134969, p2 = 13276/131896, nsim = 75, alternative = "greater")),
                 generic_turf_top5_power = unlist(pmap(args_n, power.fisher.test, p1 = 7370/134969, p2 = 7146/131896, nsim = 75, alternative = "greater")),
                 generic_turf_top1_power = unlist(pmap(args_n, power.fisher.test, p1 = 1353/134969, p2 = 1341/131896, nsim = 75, alternative = "greater")))

```

```{r reshape generic turf df}
generic_turf_long_df <-
  generic_turf_df %>%
  pivot_longer(!sample_size, names_to = "annotation", values_to = "power")
```

```{r plot generic turf power}
ggplot(generic_turf_long_df, aes(x=sample_size/70, y=power)) + 
  #geom_point(aes(color=annotation)) + 
  geom_smooth(method = "loess", se = FALSE, aes(color=annotation), span = .9) +
  geom_vline(xintercept = 1917) + #number of observed proband mutations
  geom_hline(yintercept = .8, linetype = "dashed")
```

```{r combine generic and brain turf dataframes}
combined_turf_long_df <- rbind(generic_turf_long_df, brain_turf_long_df)
```

```{r plot generic and brain turf}
ggplot(combined_turf_long_df, aes(x=sample_size, y=power)) + 
  #geom_point(aes(color=annotation)) + 
  geom_smooth(method = "loess", se = FALSE, aes(color=annotation), span = .9) +
  geom_hline(yintercept = .8, linetype = "dashed")
```


### RegulomeDB power

```{r calculate regdb power}
# Can't calculate regdb 3's because the sibling proportion is greater
args_n <- list(n1 = seq(1, 5000000, 100000), n2 = seq(1, 5000000, 100000))

regdb_df = data.frame(sample_size = seq(1, 5000000, 100000),
                 regdb_all2s_power = unlist(pmap(args_n, power.fisher.test, p1 = 5446/134969, p2 = 5196/131896, nsim = 75, alternative = "greater")),
                 regdb_2a_power = unlist(pmap(args_n, power.fisher.test, p1 = 524/134969, p2 = 493/131896, nsim = 75, alternative = "greater")),
                 regdb_2b_power = unlist(pmap(args_n, power.fisher.test, p1 = 4473/134969, p2 = 4304/131896, nsim = 75, alternative = "greater")),
                 regdb_2c_power = unlist(pmap(args_n, power.fisher.test, p1 = 449/134969, p2 = 399/131896, nsim = 75, alternative = "greater")))
```

```{r reshape regdb df}
regdb_long_df <-
  regdb_df %>%
  pivot_longer(!sample_size, names_to = "annotation", values_to = "power")
```

```{r plot regdb power}
ggplot(regdb_long_df, aes(x=sample_size/70, y=power)) + 
  #geom_point(aes(color=annotation)) + 
  geom_smooth(method = "loess", se = FALSE, aes(color=annotation)) +
  geom_vline(xintercept = 1917) + #number of observed proband mutations
  geom_vline(xintercept = 43000) + #number of observed proband mutations
  geom_curve(aes(x = 20000, y = .15, xend = 2200, yend = .03),arrow = arrow(length = unit(0.03, "npc")), curvature = .1) +
  annotate('text', x = 28000, y = .15, label = 'This Study' ) +
  geom_segment(aes(x = 55000, y =.3 , xend = 43500, yend = .4),
                  arrow = arrow(length = unit(0.5, "cm"))) +
  annotate('text', x = 58000, y = .27, label = 'SPARK' ) +
  geom_hline(yintercept = .8, linetype = "dashed")
```
                

### VEP power
```{r calculate vep power}
args_n <- list(n1 = seq(1, 5000000, 100000), n2 = seq(1, 5000000, 100000))

vep_df = data.frame(sample_size = seq(1, 5000000, 100000),
                 missense_power = unlist(pmap(args_n, power.fisher.test, p1 = 1572/134969, p2 = 1480/131896, nsim = 75, alternative = "greater")),
                 regulatory_region_power = unlist(pmap(args_n, power.fisher.test, p1 = 2356/134969, p2 = 2185/131896, nsim = 75, alternative = "greater")),
                 splice_region_power = unlist(pmap(args_n, power.fisher.test, p1 = 187/134969, p2 = 167/131896, nsim = 75, alternative = "greater")),
                 stop_gain_power = unlist(pmap(args_n, power.fisher.test, p1 = 120/134969, p2 = 62/131896, nsim = 75, alternative = "greater")))
```

```{r reshape vep df}
vep_long_df <-
  vep_df %>%
  pivot_longer(!sample_size, names_to = "annotation", values_to = "power")
```

```{r plot vep power}
ggplot(vep_long_df, aes(x=sample_size/70, y=power)) + 
  #geom_point(aes(color=annotation)) + 
  geom_smooth(method = "loess", se = FALSE, aes(color=annotation), span = .6) +
  geom_hline(yintercept = .8, linetype = "dashed") +
  geom_vline(xintercept = 1917) + #number of observed proband mutations
  geom_curve(aes(x = 20000, y = .15, xend = 1917, yend = .05),arrow = arrow(length = unit(0.03, "npc")), curvature = .1) +
  annotate('text', x = 28000, y = .15, label = 'This Study') +
  scale_x_continuous(name = "Number of Families") +
  scale_y_continuous(name = "Power", breaks = c(0,.2,.4,.6,.8,1), limits = c(0,1)) + #the fitted curve would go above 1 without setting this limit
  theme_minimal(base_size = 12) +
  theme(axis.title.x = element_text(margin = margin(t=12, r=0, b=0, l=0)), #space between axis labels and title
        axis.title.y = element_text(margin = margin(t=0, r=10, b=0, l=0)))
```

```{r combine all annotation dataframes}
all_annots_long_df <- rbind(brain_promoter_long_df, combined_turf_long_df, regdb_long_df, vep_long_df)
```

```{r plot combination of annotations}
selected_annotations <- c("fetalbrain_promoter_power", 
                          "brain_turf_top5_power",
                          "brain_turf_top1_power",  #generic top 1% has larger sib proportion
                          "generic_turf_top5_power",
                          "regdb_all2s_power",
                          "stop_gain_power",
                          "regulatory_region_power")

ggplot(all_annots_long_df[all_annots_long_df$annotation %in% selected_annotations,], 
       aes(x=sample_size/70, y=power)) + 
  #geom_point(aes(color=annotation)) + 
  geom_smooth(method = "loess", se = FALSE, aes(color=annotation)) +
  geom_hline(yintercept = .8, linetype = "dashed") +
  geom_vline(xintercept = 1917)  #number of observed proband mutations

```


