---
title: "pwr package"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pwr)
library(tidyverse)
library(here)
```

# The possible parameters are
# h = effect size, or "d" when using cohen's d (difference between means divided by pooled S.dev.)
# n = sample size
# sig.level = significance level (Type I error probability; default is .05)
# power = desired power (1 - beta/TypeII error prob.)
# alternative = alternative hypothesis, "two.sided" is default

# Leave out the parameter you wish to solve for. 

```{r one-sample proportion test}
# how many times should a coin be flipped to have .80 power of correctly 
# rejecting null of .5 if a coin is loaded to land heads 75% of the time
p.out <-
  pwr.p.test(h = ES.h(p1 =.75, p2 = .5), #(ES.h(p1, p2)) for effect size calculation for proportions
          sig.level = .05,
          power = .80,
          alternative = "greater") #says we think the alternative is greater than the null

# Plot a power curve
plot(p.out)

```


```{r use range of sample sizes}
n <- seq(10,100,10)

p.out <- # creates a list object
  pwr.p.test(h = .5, # effect size
             sig.level = .05,
             n = n)
data.frame(n, power = sprintf("%.2f%%", p.out$power * 100))

# Extract specific quantities
pwr.p.test(h = .5, n = n, sig.level = .05)$power

plot(p.out)
```

```{r two-sample t test for means}
# Is there a difference between what male and female students pay at a coffee shop
# observe 30 male and 30 female students and calculate the mean purchase price for each gender.
 
# How powerful does the experiment have to be to detect a "medium" effect in either
# direction with a significance level of .05?

# Use the built-in conventional cohen effect size function (small, medium, or large)
cohen.ES(test = "t", size = "medium") # this corresponds to an effect size of .5

pwr.t.test(n = 30, d = .5, sig.level = .05) # n is the number in EACH group, d is cohen's effect size

# This tells us our power is .47
# How many students should be observed for a test with 80% power?

pwr.t.test(d = 0.5, power = 0.80, sig.level = 0.05)
#  this shows that n should be 63
```

```{r paired t test}
##### Example 
# 40-yard dash time for 24 runners are measured before and after training
# We want to see if they are faster after training
# how powerful is a test to detect a difference of .08 seconds with .05 significance, if
#     we assume standard deviation of the differences is .25 seconds

# We need to calculate cohen's d when we have a specific effect amount.  
# d = difference in means/ standard deviation

pwr.t.test(n = 24, d = .08/.25,
           type = "paired", alternative = "greater")
# power is only .45 in this case

# How many runners should be sampled for 80% power?
pwr.t.test(d = 0.08 / 0.25, power = 0.8,
           type = "paired", alternative = "greater")
# n should be 62

# If the standard deviation WITHIN pairs is estimated instead of for the difference in pairs:
# we need to multiply the estimated standard deviaion by sqrt(2).  
#So, if we estimate standard deviaion of each runner's 40 yard dash to be .10 seconds

pwr.t.test(d = 0.08 / (0.1 * sqrt(2)), power = 0.8,
           type = "paired", alternative = "greater")
# sample size now needed to detect a difference of .08 seconds is 21```
```

```{r fetal brain promoter power}
power.prop.test(p1=(1806/134969), p2=(1732/131896),
                sig.level = .05,
                n = 134969,
                #power = .80,
                alternative = "one.sided")
```

```{r proximity to tss promoter power}
power.prop.test(p1=1517/134969, p2 = 1416/131896,
                n = 134969,
                sig.level = .05,
                #power = .80,
                alternative = "one.sided")

```




```{r 2 proportions, unequal n for fetal brain promoter}
pwr.2p2n.test(h=ES.h(p1 = (1806/134969), p2 = (1732/131896)), 
              sig.level = .05,
              n1 = 134969, n2 = 131896, 
              alternative = "greater")
```
```{r 2 proportion equal n for fetal brain promoter}
pwr.2p.test(h=ES.h(p1 = (1806/134969), p2 = (1732/131896)),
            sig.level = .05,
            power = .80)

# n =3303948 (mutations per child), about 47k quads
```
```{r 2 proportion equl n FB promoter, changing n}
pwr.2p.test(h=ES.h(p1 = (1806/134969), p2 = (1732/131896)),
            sig.level = .05,
            n = c(2000*70, 20000*70, 40000*70, 60000*70)) #families times average mutations per child

```




```{r 2 proportion equl n FB promoter, changing n , save results to data frame}
n <- seq(1, 5000000, 100000)  # Increments of 100000/70 mutations ~1500 families
p.out <- pwr.2p.test(h=ES.h(p1 = (1806/134969), p2 = (1732/131896)),
            sig.level = .05,
            n = n)

p.out.df <- data.frame(n, power=p.out$power) #extract power from results with $

```


```{r plot from pwr results dataframe}
ggplot(p.out.df, aes(x=n/70, y=power)) + 
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) + 
  geom_hline(yintercept = .8, linetype = "dashed") +
  geom_vline(xintercept = 1917, linetype = "dashed") +
  scale_y_continuous(name = "Power", breaks = c(0,.2,.4,.6,.8,1), limits = c(0,1))

```

```{r 2 proportion equl n proximity to tss promoter, changing n , save results to data frame}
n <- seq(1, 5000000, 50000)  # Increments of 100000/70 mutations ~1500 families
p.out <- pwr.2p.test(h=ES.h(p1 = 1517/134969, p2 = 1416/131896),
            sig.level = .05,
            n = n)

p.out.proxtoTSS.df <- data.frame(n, power=p.out$power) #extract power from results with $
```

```{r plot from proxtoTSS pwr results dataframe}
ggplot(p.out.proxtoTSS.df, aes(x=n/70, y=power)) + 
  #geom_point() +
  geom_line(color = "blue") +
  #geom_smooth(method = "loess", se = FALSE) + 
  geom_hline(yintercept = .8, linetype = "dashed") +
  geom_vline(xintercept = 1917, linetype = "dotted") +
  scale_y_continuous(name = "Power", breaks = c(0,.2,.4,.6,.8,1), limits = c(0,1))

```

```{r TURF power}
n <- seq(1, 4800000, 10000)  # Increments of 100000/70 mutations ~1500 families
fbPromoter.out <- pwr.2p.test(h=ES.h(p1 = (1806/134969), p2 = (1732/131896)),
            sig.level = .05,
            n = n)

genericTURF.out <- pwr.2p.test(h=ES.h(p1 = 7370/134969, p2 = 7146/131896),
            sig.level = .05,
            n = n)

brainTURF.out <- pwr.2p.test(h=ES.h(p1 = 6828/134969, p2 = 6540/131896),
            sig.level = .05,
            n = n)

#stopgain.out <- pwr.2p.test(h=ES.h(p1 = 120/134969, p2 = 62/131896),
#            sig.level = .05,
#            n = n)

missense.out <- pwr.2p.test(h=ES.h(p1 = (1572/134969), p2 = (1480/131896)),
            sig.level = .05,
            n = n)

coding_vep_highimpact.out <- pwr.2p.test(h=ES.h(p1 = (126/134969), p2 = (74/131896)),    #(604/134969), p2 = (467/131896)
            sig.level = .05,
            n = n)

missense.out_df <- data.frame(n, missensepower=missense.out$power)

#stopgain.out_df <- data.frame(n, stopgainpower=stopgain.out$power)

fbPromoter.out_df <- data.frame(n, fbPromoterpower=fbPromoter.out$power)

genericTURF.out_df <- data.frame(n, genericTURFpower=genericTURF.out$power)

brainTURF.out_df <- data.frame(n, brainTURFpower=brainTURF.out$power)

coding_vep_highimpact.out_df <- data.frame(n, codingpower=coding_vep_highimpact.out$power)


mergedAnno.out_df <- join_all(list(genericTURF.out_df, brainTURF.out_df, fbPromoter.out_df, missense.out_df,  # stopgain.out_df, missense.out_df,
                                   coding_vep_highimpact.out_df))

mergedAnno_long_df <-
  mergedAnno.out_df %>%
  pivot_longer(!n, names_to = "annotation", values_to = "power")

```


```{r}
ggplot(mergedAnno_long_df, aes(x=n/70, y=power, color = fct_reorder2(annotation, n, power))) + #order factors so lines match order in legent
  #geom_point() +
  geom_line(size = 1.3) +
  #geom_smooth(method = "loess", se = FALSE, aes(color = annotation)) + 
  #geom_hline(yintercept = .8, linetype = "dashed") +
  geom_vline(xintercept = 1917, linetype = "dashed") +
  geom_curve(aes(x = 14000, y = .15, xend = 2300, yend = .03),
             arrow = arrow(length = unit(0.03, "npc")), color = "black", curvature = .1) +
  annotate('text', x = 19000, y = .15, label = 'This Study') +
  #geom_segment(
    #x = 12000, y = .06,
    #xend = 2300, yend = .02,
    #lineend = "butt", # See available arrow types in example above
    #linejoin = "round",
    #size = .75, 
    #arrow = arrow(length = unit(0.12, "inches"), type = "closed"),
    #color = "black") +
  scale_y_continuous(breaks = c(0,.2,.4,.6,.8,1), limits = c(0,1)) +
  theme_minimal(base_size = 12) +
  theme(axis.title.x = element_text(margin = margin(t=6, r=0, b=0, l=0)), #space between axis labels and title
        axis.title.y = element_text(margin = margin(t=0, r=6, b=0, l=0)),
        legend.position = c(0.80, 0.30),
        legend.background = element_rect(fill = "white", color = "black"),
        legend.key.size = unit(.5, 'cm')) + # change legend box size
  scale_color_discrete(name = "Genomic Annotation", labels = c(#"stop-gain", 
    "high-impact coding", "all missense","brain-tissue TURF top 5%", "fetal brain promoter", "generic TURF top 5%")) + # change legend label names
  labs(
    x = "Number of quads",
    y = "Power")

```
```{r}
ggsave(filename = here::here("figures", "annotation_power.png"), width = 1900, height = 1200, units = "px")
```


